---
title: 'Project 1: Linear Programming'
author: "Brooks Beckelman"
date: "February 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(lpSolve)
```

## Function Inputs 

```{r}
# Liabilities (in dollars)
l = c(12000,18000,20000,20000,16000,15000,12000,10000)

# Bond prices (in dollars)
p = c(102,99,101,98,98,104,100,101,102,94)

# Coupons (in dollars)
coup = c(5,3.5,5,3.5,4,9,6,8,9,7)

# Maturities (in years)
m = c(1,2,2,3,4,5,5,6,7,8)
```

## 1. Formulate portfolio construction problem as a linear program

Decision Variables: amount of each bond (1-10) to purchase (x = x1,...,x10)

Objective: minimize the total cost of the bonds (p*x)

Constraints: cash flow inward from coupon and maturity payments must equal cash flow outward to liabilities for each year (1-8)

```{r}
# Objective
c = p

# Constraints
A = matrix(0,length(l),length(c))

for (i in seq_along(coup)){
  ## Fill matrix with coupon payments
  A[1:m[i]-1,i] = coup[i]
  ## Fill matrix with maturity payments and final coupon payments
  A[m[i],i] = 100 + coup[i]
}

## Direction Vector
dir = rep("=",length(l))

## Constraints Vector 
b = l
```

## 2. Test Case

```{r}
# Solve LP
s = lp("min",c,A,dir,b)
sol = s$solution

# Print Optimal Solution
for (i in seq_along(sol)){
  cat("Bond",i,"\t:\t",sol[i],"\n")
}
```

## 3. Portfolio Function

The following function (dedicate_g4) can construct a portfolio for any set of liabilities and bonds. The function takes four inputs: P, C, M, and L. P is the vector containing the prices of the bonds. C is the vector containing the coupon payments for the bonds. M is the vector containing the maturity (in years) for the bonds. Finally, L is the vector of non-negative liabilities for each year. The function outputs the optimal number of each bond to purchase in order to minimize the amount paid for the bonds up front.

```{r}
dedicate_g4 <- function(P,C,M,L){
  # Set objective
  c = P
  
  # Constraints
  ## Initialize constraints matrix with zeros
  A = matrix(0,length(L), length(C))
  
  for (i in seq_along(C)){
    ## Fill matrix with coupon payments
    A[1:M[i]-1,i] = C[i]
    
    ## Fill matrix with maturity payments + final coupon payment
    A[M[i],i] = 100 + C[i]
  }
  
  ## Direction Vector
  dir = rep("=",length(L))
  
  ## Constraints Vector
  b = L
  
  # Solution
  s = lp("min",c,A,dir,b,compute.sens=1)
  return(s)
}
```

Let's test the function using our test case from problem 2.

```{r}
# Function Inputs
## Liabilities (in dollars)
l = c(12000,18000,20000,20000,16000,15000,12000,10000)

## Bond prices (in dollars)
p = c(102,99,101,98,98,104,100,101,102,94)

## Coupons (in dollars)
coup = c(5,3.5,5,3.5,4,9,6,8,9,7)

## Maturities (in years)
m = c(1,2,2,3,4,5,5,6,7,8)

# Run function
s = dedicate_g4(p,coup,m,l)
sol = s$solution

# Print Optimal Solution
for (i in seq_along(sol)){
  cat("Bond",i,"\t:\t",sol[i],"\n")
}
```

The function outputs the same result that we saw from building the portfolio manually. 

## 4. Construct a dedicated portfolio

Bond information was collected from the Wall Street Journal (http://online.wsj.com/mdc/public/page/2_3020-treasury.html#treasuryB) and stored as a csv file named "TreasuryQuotes_wsj.csv".

```{r}
# Read in bond information and store as data frame
setwd(setwd("C:/Users/brook/Desktop/MSBA_Spring2017/Optimization/Assignments/Project1"))
bonds = read.csv("TreasuryQuotes_wsj.csv")
```

Using the 'bonds' data frame, we can now create our input vectors.

```{r}
# Limit bonds dataframe to only those that mature/pay coupons at the end of June/December
dates = c("6/30/2017","12/31/2017","6/30/2018","12/31/2018","6/30/2019","12/31/2019",
          "6/30/2020","12/31/2020","6/30/2021","12/31/2021","6/30/2022","12/31/2022")

bonds = bonds[bonds$Maturity %in% dates,]

# Price Vector
P = bonds$Asked

# Coupon Vector
C = bonds$Coupon/2

# Maturities Vector (in periods/half-years)
## Assign each date to period number
periods = seq_along(dates)
names(periods) = dates

## Initialize maturities vector with zeros
M = rep(0,length(bonds$Maturity))

## Add period number corresponding to each maturity date to maturities vector
for (i in seq_along(M)){
  date = toString(bonds$Maturity[i])
  period = periods[date]
  
  M[i] = period
}

# Non-negative liabilities vector (in dollars)
L = 1000000*c(9,9,10,10,6,6,9,9,10,10,5,3)
```

Finally, we can solve for the optimal amount of each bond to purchase using the function we defined previously.

```{r}
options("scipen"=100, "digits"=4)
# Run function
s = dedicate_g4(P,C,M,L)

optimal_solution = data.frame(bonds$Maturity,bonds$Coupon,bonds$Asked,s$solution)
names(optimal_solution) = c("Maturity","Coupon","Price","Amt. to Purchase")
optimal_solution
```

### Sensitivity Analysis
Next we can evaluate the sensitivity of this model to the liability constraints.

```{r}
optimal_solution_sens = data.frame(dates,L,s$duals[0:length(L)],s$duals.from[0:length(L)],s$duals.to[0:length(L)])
names(optimal_solution_sens) = c("Date","Liability","Duals","Duals From","Duals To")

optimal_solution_sens
```

```{r}
plot(periods, optimal_solution_sens$Duals, main="Liability Constraint Sensitivities Over Time", xlab="Liability Date", ylab="Duals", type='o', col='blue', xaxt='n')
axis(1, at=1:12, labels=dates[1:12])
```

The duals (i.e. sensitivity of the constraint for each half-year) can be interperted as the time value of money. In other words,  1 dollar on 12/31/2022 is worth 88 cents today. The Duals From & To are the range of liabilities for that date for which the duals/sensitivity applies.