---
title: 'Project 1: Linear Programming'
author: "Brooks Beckelman"
date: "February 4, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(lpSolve)
library(XML)
```

## Function Inputs 

```{r}
# Liabilities (in dollars)
l = c(12000,18000,20000,20000,16000,15000,12000,10000)

# Bond prices (in dollars)
p = c(102,99,101,98,98,104,100,101,102,94)

# Coupons (in dollars)
coup = c(5,3.5,5,3.5,4,9,6,8,9,7)

# Maturities (in years)
m = c(1,2,2,3,4,5,5,6,7,8)
```

## 1. Formulate portfolio construction problem as a linear program

Decision Variables: amount of each bond (1-10) to purchase (x = x1,...,x10)

Objective: minimize the total cost of the bonds (p*x)

Constraints: cash flow inward from coupon and maturity payments must equal cash flow outward to liabilities for each year (1-8)

```{r}
# Objective
c = p

# Constraints
A = matrix(0,length(l),length(c))

for (i in seq_along(coup)){
  ## Fill matrix with coupon payments
  A[1:m[i]-1,i] = coup[i]
  ## Fill matrix with maturity payments and final coupon payments
  A[m[i],i] = 100 + coup[i]
}

## Direction Vector
dir = rep("=",length(l))

## Constraints Vector 
b = l
```

## 2. Test Case

```{r}
# Solve LP
s = lp("min",c,A,dir,b)
sol = s$solution

# Print Optimal Solution
for (i in seq_along(sol)){
  cat("Bond",i,"\t:\t",sol[i],"\n")
}
```

## 3. Portfolio Function

The following function (dedicate_g4) can construct a portfolio for any set of liabilities and bonds. The function takes four inputs: P, C, M, and L. P is the vector containing the prices of the bonds. C is the vector containing the coupon payments for the bonds. M is the vector containing the maturity (in years) for the bonds. Finally, L is the vector of non-negative liabilities for each year. The function outputs the optimal number of each bond to purchase in order to minimize the amount paid for the bonds up front.

```{r}
dedicate_g4 <- function(P,C,M,L){
  # Set objective
  c = P
  
  # Constraints
  ## Initialize constraints matrix with zeros
  A = matrix(0,length(L), length(C))
  
  for (i in seq_along(C)){
    ## Fill matrix with coupon payments
    A[1:M[i]-1,i] = C[i]
    
    ## Fill matrix with maturity payments + final coupon payment
    A[M[i],i] = 100 + C[i]
  }
  
  ## Direction Vector
  dir = rep("=",length(L))
  
  ## Constraints Vector
  b = l
  
  # Solution
  s = lp("min",c,A,dir,b)
  return(s$solution)
}
```

Let's test the function using our test case from problem 2.

```{r}
# Function Inputs
## Liabilities (in dollars)
l = c(12000,18000,20000,20000,16000,15000,12000,10000)

## Bond prices (in dollars)
p = c(102,99,101,98,98,104,100,101,102,94)

## Coupons (in dollars)
coup = c(5,3.5,5,3.5,4,9,6,8,9,7)

## Maturities (in years)
m = c(1,2,2,3,4,5,5,6,7,8)

# Run function
sol = dedicate_g4(p,coup,m,l)

# Print Optimal Solution
for (i in seq_along(sol)){
  cat("Bond",i,"\t:\t",sol[i],"\n")
}
```

The function outputs the same result that we saw from building the portfolio manually. 

## 4. Construct a dedicated portfolio

```{r}
# Non-negative liabilities vector (in dollars)
L = 1000000*c(9,9,10,10,6,6,9,9,10,10,5,3)
```

